---
title: "Knowledge Graph Hate Speech Analysis Report"
subtitle: "Machine vs Human Classification Comparison"
author: "Randy Howk"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)

# Load required libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(scales)
library(viridis)
library(patchwork)
library(plotly)

# Network visualization
library(igraph)
library(ggraph)
library(tidygraph)

# Dimensionality reduction
library(umap)

# Statistical testing
library(broom)

# Read configuration
library(jsonlite)

# Set theme
theme_set(theme_minimal(base_size = 12))

```
## Executive Summary
```{r executive-summary}
# Load metadata
metadata <- fromJSON("output/r_metadata.json")

# Load main datasets
incidents <- read_csv("output/structured_incidents.csv", show_col_types = FALSE)
category_performance <- read_csv("output/category_performance.csv", show_col_types = FALSE)
confusion_matrix <- read_csv("output/confusion_matrix_long.csv", show_col_types = FALSE)
confidence_analysis <- read_csv("output/confidence_analysis.csv", show_col_types = FALSE)

# Load evaluation metrics
eval_metrics <- fromJSON("results/evaluation_metrics.json")

# Display evaluation report
eval_report <- readLines("results/evaluation_report.txt")
```

This report presents the results of a knowledge graph-enhanced classification system for hate incident terminology, comparing machine learning predictions against human expert classifications.

Key Findings

```{r key-findings}
# Calculate key metrics
total_incidents <- nrow(incidents)
overall_accuracy <- eval_metrics$accuracy
f1_score <- eval_metrics$f1_weighted
n_categories <- metadata$statistics$n_categories

# Create summary table
summary_stats <- tibble(
  Metric = c(
    "Total Incidents Analyzed",
    "Number of Categories",
    "Overall Accuracy",
    "Weighted F1-Score",
    "Macro F1-Score",
    "Precision (Weighted)",
    "Recall (Weighted)"
  ),
  Value = c(
    format(total_incidents, big.mark = ","),
    n_categories,
    percent(overall_accuracy, accuracy = 0.01),
    round(f1_score, 3),
    round(eval_metrics$f1_macro, 3),
    round(eval_metrics$precision_weighted, 3),
    round(eval_metrics$recall_weighted, 3)
  )
)

kable(summary_stats, 
      caption = "Overall Performance Summary",
      align = c('l', 'r')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

1. Performance Metrics
1.1 Overall Accuracy

```{r overall-accuracy}
# Create accuracy visualization
accuracy_data <- tibble(
  Model = c("Knowledge Graph Enhanced", "Random Baseline", "Target"),
  Accuracy = c(overall_accuracy, 1/n_categories, 0.85)
)

ggplot(accuracy_data, aes(x = Model, y = Accuracy, fill = Model)) +
  geom_col() +
  geom_text(aes(label = percent(Accuracy, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  geom_hline(yintercept = 0.85, linetype = "dashed", color = "red", size = 1) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_fill_viridis_d(option = "C") +
  labs(
    title = "Model Performance vs Baselines",
    subtitle = "Dashed line indicates target performance (85%)",
    x = NULL,
    y = "Accuracy"
  ) +
  theme(legend.position = "none")
```

1.2 Performance by Confidence Level

```{r performance-by-confidence-level}
# Analyze accuracy by confidence level
confidence_perf <- incidents %>%
  filter(!is.na(confidence_level)) %>%
  group_by(confidence_level) %>%
  summarise(
    accuracy = mean(is_correct, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  ) %>%
  mutate(confidence_level = factor(confidence_level, levels = c("Low", "Medium", "High")))

# Visualization
ggplot(confidence_perf, aes(x = confidence_level, y = accuracy, fill = confidence_level)) +
  geom_col() +
  geom_text(aes(label = paste0(percent(accuracy, accuracy = 0.1), "\n(n=", count, ")")), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_fill_manual(values = c("Low" = "#d62728", "Medium" = "#ff7f0e", "High" = "#2ca02c")) +
  labs(
    title = "Accuracy by Confidence Level",
    subtitle = "Higher confidence predictions are more accurate",
    x = "Confidence Level",
    y = "Accuracy"
  ) +
  theme(legend.position = "none")

# Create detailed table
confidence_perf %>%
  mutate(accuracy = percent(accuracy, accuracy = 0.01)) %>%
  kable(
    caption = "Detailed Performance by Confidence Level",
    col.names = c("Confidence Level", "Accuracy", "Number of Incidents")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
2. Category-Level Analysis
2.1 Top Performing Categories
```{r some-name-6}
# Get top 15 categories by F1 score
top_categories <- category_performance %>%
  arrange(desc(f1_score)) %>%
  head(15)

ggplot(top_categories, aes(x = reorder(category, f1_score), y = f1_score)) +
  geom_col(aes(fill = f1_score)) +
  geom_text(aes(label = round(f1_score, 2)), hjust = -0.1, size = 3) +
  coord_flip() +
  scale_fill_viridis_c(option = "D") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Top 15 Best Performing Categories",
    subtitle = "Ranked by F1-Score",
    x = NULL,
    y = "F1-Score"
  ) +
  theme(legend.position = "none")
```

2.2 Category Performance Distribution

```{r category-performance-distribution}
# Precision vs Recall scatter
ggplot(category_performance, aes(x = recall, y = precision, size = support, color = f1_score)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_viridis_c(option = "C") +
  scale_size_continuous(range = c(2, 10)) +
  labs(
    title = "Precision vs Recall by Category",
    subtitle = "Point size represents number of samples, color represents F1-score",
    x = "Recall",
    y = "Precision",
    color = "F1-Score",
    size = "Support"
  ) +
  theme(legend.position = "right")
```

2.3 Performance Metrics Heatmap

```{r performance-metrics-heatmap}
# Create heatmap of top categories
top_20_cats <- category_performance %>%
  arrange(desc(f1_score)) %>%
  head(20)

# Reshape for heatmap
metrics_long <- top_20_cats %>%
  select(category, precision, recall, f1_score) %>%
  pivot_longer(cols = c(precision, recall, f1_score),
               names_to = "metric",
               values_to = "value")

ggplot(metrics_long, aes(x = metric, y = reorder(category, value), fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(value, 2)), size = 3) +
  scale_fill_viridis_c(option = "A") +
  labs(
    title = "Performance Metrics Heatmap",
    subtitle = "Top 20 categories by F1-score",
    x = "Metric",
    y = NULL,
    fill = "Score"
  ) +
  theme(axis.text.x = element_text(angle = 0))
```

3. Confusion Matrix Analysis
3.1 Confusion Matrix Visualization

```{r confusion-matrix-visualization}
# Get top categories for confusion matrix (limit for readability)
top_cats <- incidents %>%
  count(manual_term, sort = TRUE) %>%
  head(15) %>%
  pull(manual_term)

# Filter confusion matrix
cm_filtered <- confusion_matrix %>%
  filter(true_category %in% top_cats & predicted_category %in% top_cats)

# Create heatmap
ggplot(cm_filtered, aes(x = predicted_category, y = true_category, fill = count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = count), size = 3) +
  scale_fill_viridis_c(option = "B", trans = "log1p") +
  labs(
    title = "Confusion Matrix: Top 15 Categories",
    subtitle = "Rows = Human Classification, Columns = Machine Prediction",
    x = "Machine Predicted Category",
    y = "Human Classified Category (Ground Truth)",
    fill = "Count\n(log scale)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9)
  )
```

3.2 Common Misclassifications

```{r common-misclassifications, error=TRUE}
# Identify most common misclassifications
misclassifications <- confusion_matrix %>%
  filter(true_category != predicted_category) %>%
  arrange(desc(count)) %>%
  head(15)

ggplot(misclassifications, 
       aes(x = reorder(paste(true_category, "→", predicted_category), count), 
           y = count)) +
  if (nrow(top_misclass) > 0) {
  geom_col(fill = "#e74c3c") +
  } else {
    print("No misclassifications to display")
  }
  geom_text(aes(label = count), hjust = -0.1, size = 3) +
  coord_flip() +
  labs(
    title = "Top 15 Most Common Misclassifications",
    subtitle = "Format: True Category → Predicted Category",
    x = NULL,
    y = "Number of Misclassifications"
  )
```

4. Embedding Analysis
4.1 UMAP Visualization of Incident Embeddings

```{r imap-visualization-of-incident-embeddings, error=TRUE}
# Load embeddings
incident_embeddings <- read_csv("output/incident_embeddings_for_r.csv", show_col_types = FALSE)

# Prepare data for UMAP (sample if too large)
set.seed(42)
sample_size <- min(2000, nrow(incident_embeddings))
sample_indices <- sample(1:nrow(incident_embeddings), sample_size)

embedding_matrix <- incident_embeddings %>%
  select(starts_with("dim_")) %>%
  slice(sample_indices) %>%
  as.matrix()

# Run UMAP
umap_config <- umap.defaults
umap_config$n_neighbors <- 15
umap_config$min_dist <- 0.1

umap_result <- umap(embedding_matrix, config = umap_config)

# Prepare data for plotting
umap_df <- tibble(
  umap1 = umap_result$layout[, 1],
  umap2 = umap_result$layout[, 2],
  incident_id = incident_embeddings$incident_id[sample_indices]
) %>%
  tryCatch({
  left_join(incidents %>% select(incident_id, manual_term, is_correct, confidence),
            by = "incident_id")
  }, error = function(e) { print(paste("Join error:", e$message)); tibble() })
            by = "incident_id")

# Plot colored by category
p1 <- ggplot(umap_df, aes(x = umap1, y = umap2, color = manual_term)) +
  geom_point(alpha = 0.6, size = 1) +
  labs(
    title = "UMAP Projection of Incident Embeddings",
    subtitle = "Colored by True Category",
    x = "UMAP 1",
    y = "UMAP 2",
    color = "Category"
  ) +
  theme(legend.position = "none")  # Too many categories

# Plot colored by correctness
p2 <- ggplot(umap_df, aes(x = umap1, y = umap2, color = factor(is_correct))) +
  geom_point(alpha = 0.6, size = 1) +
  scale_color_manual(
    values = c("0" = "#e74c3c", "1" = "#2ecc71"),
    labels = c("0" = "Incorrect", "1" = "Correct")
  ) +
  labs(
    title = "Prediction Correctness in Embedding Space",
    x = "UMAP 1",
    y = "UMAP 2",
    color = "Prediction"
  ) +
  theme(legend.position = "right")

p1 / p2
```

4.2 Glossary Term Embeddings
```{r glossary-term-embeddings, error=TRUE}
# Load glossary embeddings
glossary_embeddings <- read_csv("output/glossary_embeddings_for_r.csv", show_col_types = FALSE)

# Run UMAP on glossary
glossary_matrix <- glossary_embeddings %>%
  select(starts_with("dim_")) %>%
  as.matrix()

glossary_umap <- umap(glossary_matrix, config = umap_config)

# Prepare plotting data
glossary_umap_df <- tibble(
  umap1 = glossary_umap$layout[, 1],
  umap2 = glossary_umap$layout[, 2],
  term = glossary_embeddings$term,
  category = glossary_embeddings$category
)

# Plot
ggplot(glossary_umap_df, aes(x = umap1, y = umap2, color = category)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_viridis_d(option = "D") +
  labs(
    title = "UMAP Projection of Glossary Term Embeddings",
    subtitle = "Terms cluster by semantic category",
    x = "UMAP 1",
    y = "UMAP 2",
    color = "Category"
  ) +
  theme(legend.position = "right")
```

5. Knowledge Graph Visualization
5.1 Term Co-occurrence Network

```{r term-co-occurrence-network}
# Load graph data
tryCatch({
  graph_nodes <- read_csv("output/graph_nodes.csv", show_col_types = FALSE)
  graph_edges <- read_csv("output/graph_edges.csv", show_col_types = FALSE)
  
  # Create igraph object
  g <- graph_from_data_frame(
    d = graph_edges %>% select(source, target, type, score),
    vertices = graph_nodes,
    directed = TRUE
  )
  
  # Convert to tidygraph for ggraph
  tg <- as_tbl_graph(g)
  
  # Calculate centrality
  tg <- tg %>%
    activate(nodes) %>%
    mutate(
      degree = centrality_degree(),
      betweenness = centrality_betweenness()
    )
  
  # Plot
  ggraph(tg, layout = "fr") +
    geom_edge_link(aes(edge_alpha = score), 
                   arrow = arrow(length = unit(2, 'mm')),
                   end_cap = circle(3, 'mm'),
                   color = "gray70") +
    geom_node_point(aes(size = degree, color = category), alpha = 0.8) +
    geom_node_text(aes(label = name, filter = degree > quantile(degree, 0.9)), 
                   repel = TRUE, size = 3) +
    scale_size_continuous(range = c(2, 10)) +
    scale_color_viridis_d(option = "C") +
    labs(
      title = "Knowledge Graph: Term Relationships",
      subtitle = "Node size represents degree centrality, labels shown for top 10% most connected terms",
      color = "Category",
      size = "Connections"
    ) +
    theme_graph() +
    theme(legend.position = "right")
  
}, error = function(e) {
  cat("Graph visualization data not available\n")
})
```

6. Temporal Analysis
```{r temporal-analysis}
# Load temporal data
tryCatch({
  temporal_data <- read_csv("output/temporal_analysis.csv", show_col_types = FALSE)
  
  # Incidents over time
  temporal_summary <- temporal_data %>%
    filter(!is.na(year_month)) %>%
    group_by(year_month) %>%
    summarise(
      count = n(),
      .groups = "drop"
    ) %>%
    mutate(year_month = as.Date(paste0(year_month, "-01")))
  
  ggplot(temporal_summary, aes(x = year_month, y = count)) +
    geom_line(color = "#3498db", size = 1) +
    geom_point(color = "#2980b9", size = 2) +
    scale_x_date(date_breaks = "3 months", date_labels = "%b %Y") +
    labs(
      title = "Incident Volume Over Time",
      x = "Month",
      y = "Number of Incidents"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
}, error = function(e) {
  cat("Temporal analysis data not available\n")
})
```

7. Geographic Distribution

```{r geographic-distribution, error=TRUE}
# Load geographic data
tryCatch({
  geo_data <- read_csv("output/geographic_analysis.csv", show_col_types = FALSE)
  
  # Top states by incident count
  top_states <- geo_data %>%
    group_by(state) %>%
    summarise(total = sum(count), .groups = "drop") %>%
    arrange(desc(total)) %>%
    head(20)
  
  ggplot(top_states, aes(x = reorder(state, total), y = total)) +
    geom_col(fill = "#9b59b6") +
    geom_text(aes(label = total), hjust = -0.1, size = 3) +
    coord_flip() +
    labs(
      title = "Top 20 States by Incident Count",
      x = NULL,
      y = "Number of Incidents"
    )
  
}, error = function(e) {
  cat("Geographic analysis data not available\n")
})
```

8. Statistical Significance Tests
```{r statistical-significance-tests, error=TRUE}
# Load statistical test results
stat_tests <- fromJSON("results/statistical_tests.json")

# Create summary table
stat_summary <- tibble(
  Test = c("McNemar's Test", "Chi-Square Test", "Paired t-Test"),
  Statistic = c(
    round(stat_tests$mcnemar$statistic, 4),
    round(stat_tests$chi_square$chi2, 4),
    round(stat_tests$t_test$t_statistic, 4)
  ),
  `P-Value` = c(
    format.pval(stat_tests$mcnemar$p_value, digits = 4),
    format.pval(stat_tests$chi_square$p_value, digits = 4),
    format.pval(stat_tests$t_test$p_value, digits = 4)
  ),
  Significant = c(
    stat_tests$mcnemar$significant,
    stat_tests$chi_square$significant,
    stat_tests$t_test$significant
  )
) %>%
  mutate(Significant = ifelse(Significant, "Yes ✓", "No"))

kable(stat_summary,
      caption = "Statistical Significance Tests (α = 0.05)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Interpretation
```{r interpretation, error=TRUE}
# Display interpretation
cat("\n")
if (stat_tests$mcnemar$significant) {
  cat("✓ McNemar's Test: The machine classifier performs significantly different from baseline.\n\n")
}

if (stat_tests$chi_square$significant) {
  cat("✓ Chi-Square Test: There is a significant association between true and predicted categories.\n\n")
}

if (stat_tests$t_test$significant) {
  cat("✓ Paired t-Test: Machine accuracy is significantly different from random guessing.\n\n")
}
```

9. Full Evaluation Report
```{r full-evaluation-report}
# Display the full evaluation report
cat(paste(eval_report, collapse = "\n"))
```
10. Conclusions and Recommendations
## Summary of Findings
Overall Performance: The knowledge graph-enhanced model achieved an accuracy of r percent(overall_accuracy, accuracy = 0.1) with a weighted F1-score of r round(f1_score, 3).

Confidence Calibration: Higher confidence predictions demonstrate significantly better accuracy, suggesting the model's confidence scores are well-calibrated.

Category Performance: Performance varies significantly across categories, with some achieving near-perfect classification while others require improvement.

Statistical Significance: The model performs significantly better than random baselines and shows meaningful patterns in the confusion matrix.

Recommendations
For Model Improvement:
Focus on categories with low F1-scores for additional training data
Investigate common misclassifications to understand systematic errors
Consider ensemble methods for low-confidence predictions
For Deployment:
Implement confidence thresholds for automated vs. human review
High confidence predictions (≥r metadata$thresholds$high_confidence) can be automated
Medium/Low confidence predictions should be flagged for human review
Future Research:
Expand the knowledge graph with additional contextual relationships
Incorporate temporal dynamics and emerging terminology
Explore multi-modal approaches (text + metadata)
Session Information
```{r some-name-19}
sessionInfo()
```
Report Generated: r Sys.time()